---
title: "BEA-all-attributes"
author: "Britta Schumacher"
date: "Last updated: `r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document:
    theme: yeti
    toc: yes
    toc_float: true
    code_folding: hide
---

```{r warning = FALSE, message = FALSE}
#source("BEA_load.R")
source("BEA_func.R")
library(tidyverse)
library(sp)
file_list <- Sys.glob("./data/*ALL_AREAS*.csv")
file_list

```

# Farm income and expenses

```{r warning = FALSE, message = FALSE}
income_expenses <- clean_BEA("./data/CAINC45__ALL_AREAS_1969_2019.csv") # farm income and expenses
unique(income_expenses$Description)
```

# GDP

## Contributions to percent change in real GDP by county & MSA

```{r warning = FALSE, message = FALSE}
perc_change_GDP <- clean_BEA("./data/CAGDP11__ALL_AREAS_2002_2019.csv") # contributions to percent change in real GDP by county & MSA
unique(perc_change_GDP$Description)
```

## GDP summary by county and MSA 

```{r warning = FALSE, message = FALSE}
summary_GDP <- clean_BEA("./data/CAGDP1__ALL_AREAS_2001_2019.csv") # GDP summary by county and MSA
unique(summary_GDP$Description)
```

## GDP in current dollars by county and MSA

```{r warning = FALSE, message = FALSE}
current_GDP <- clean_BEA("./data/CAGDP2__ALL_AREAS_2001_2019.csv") # GDP in current dollars by county and MSA
unique(current_GDP$Description)
```

## Chain-type quantity indexes for real GDP by county and MSA

```{r warning = FALSE, message = FALSE}
real_GDP <- clean_BEA("./data/CAGDP8__ALL_AREAS_2001_2019.csv") # chain-type quantity indxes for real GDP by county and MSA
unique(real_GDP$Description)
```

## Real GDP in chained dollars by county and MSA

```{r warning = FALSE, message = FALSE}
chained_GDP <- clean_BEA("./data/CAGDP9__ALL_AREAS_2001_2019.csv") # Real GDP in chained dollars by county and MSA
unique(chained_GDP$Description)
```

# Employment

## Total full-time and part-time employment by NAICS Industry

```{r warning = FALSE, message = FALSE}
employment_NAICS <- clean_BEA("./data/CAEMP25N__ALL_AREAS_2001_2019.csv") # total full-time and part-time employment by NAICS Industry
unique(employment_NAICS$Description)

```

## Total full-time and part-time employment by SIC Industry

```{r warning = FALSE, message = FALSE}
employment_SIC <- clean_BEA("./data/CAEMP25S__ALL_AREAS_1969_2000.csv") # total full-time and part-time employment by SIC Industry
unique(employment_SIC$Description)
```

# Income and earnings

## Annual personal income by county

```{r warning = FALSE, message = FALSE}
pers_income <- clean_BEA("./data/CAINC1__ALL_AREAS_1969_2019.csv") # Annual personal income by county
unique(pers_income$Description)
```

## Personal income and employment by major component by county

```{r warning = FALSE, message = FALSE}
inc_employ_mc <- clean_BEA("./data/CAINC4__ALL_AREAS_1969_2019.csv") # personal income and employment by major component by county
unique(inc_employ_mc$Description)
```

## Personal income by major component and earnings by NAICS Industry

```{r warning = FALSE, message = FALSE}
earn_naics <- clean_BEA("./data/CAINC5N__ALL_AREAS_2001_2019.csv") # personal income by major component and earnings by NAICS Industry
unique(earn_naics$Description)
```

## Personal income by major component and earnings by SIC Industry

```{r warning = FALSE, message = FALSE}
earn_sic <- clean_BEA("./data/CAINC5S__ALL_AREAS_1969_2000.csv") # personal income by major component and earnings by SIC Industry
unique(earn_sic$Description)
```

## Personal current transfer receipts

```{r warning = FALSE, message = FALSE}
trans_rec <- clean_BEA("./data/CAINC35__ALL_AREAS_1969_2019.csv") # personal current transfer receipts
unique(trans_rec$Description)
```

# Economic profile by county

```{r warning = FALSE, message = FALSE}
econ_profile <- clean_BEA("./data/CAINC30__ALL_AREAS_1969_2019.csv") # economic profile by county
unique(econ_profile$Description)
```

# Compensation of employees

## Compensation of employees by NAICS Industry

```{r warning = FALSE, message = FALSE}
comp_naics <- clean_BEA("./data/CAINC6N__ALL_AREAS_2001_2019.csv") # compensation of employees by NAICS Industry
unique(comp_naics$Description)
```

## Compensation of employees by SIC Industry

```{r warning = FALSE, message = FALSE}
comp_sic <- clean_BEA("./data/CAINC6S__ALL_AREAS_1998_2000.csv") # compensation of employees by SIC Industry
unique(comp_sic$Description)
```

# Gross flow of earnings

```{r warning = FALSE, message = FALSE}
gross_earn <- clean_BEA("./data/CAINC91__ALL_AREAS_1990_2019.csv") # gross flow of earnings
unique(gross_earn$Description)
```
