---
title: 'Data Visualizations & Stats, by SI Figure #'
author: "Britta Schumacher"
date: "Last updated: `r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document:
    theme: yeti
    toc: yes
    toc_float: true
---

# Set-up

### Load relevant packages & common data
```{r message = FALSE, warning = FALSE}
library(tidyverse)
library(ggplot2)
library(ggpubr)
library(sf)

# common data
counties <- readRDS("./data/coterm_cty_sf.RDS") 
states <- readRDS("./data/states_sf.RDS") 
states <- states %>% filter(STATE_FIPS %in% counties$STATEFP)
bea <- readRDS("./data/bea.RDS")
c <- readRDS("./data/contemp-coa.RDS")
r <- readRDS("./data/coa-race.RDS")
h <- read.csv("./data/historical-coa.csv")

# Build spatial
c_sf <- merge(counties, c, by = "GEOID")

# notes from EB
project_theme <- theme_bw() +
  theme(text = element_text(size=8))

# 
paste_noNA <- function(x,sep=", ") {
gsub(", " ,sep, toString(x[!is.na(x) & x!="" & x!="NA"] ) ) }
```


### Pull in theme for maps
```{r}
theme_map <- function(...) {
  theme_minimal() +
  theme(
    # remove all axes
    axis.line = element_blank(),
    axis.text.x = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks = element_blank(),

    # borders and margins
    panel.border = element_blank(),
    
    # titles
    legend.title = element_blank(),
    legend.text = element_text(size = 8, hjust = 0,
                               color = "black"),
    plot.title = element_text(size = 15, hjust = 0.5,
                              color = "black"),
    plot.subtitle = element_text(size = 10, hjust = 0.5,
                                 color = "black",
                                 margin = margin(b = -0.1,
                                                 t = -0.1,
                                                 l = 2,
                                                 unit = "cm"),
                                 debug = F),
    ...
  )
}
```

# Visualizations

### **SI Figure 1**: 

#### The agricultural contribution to a county’s GDP in 2019. Data: US BEA. 
```{r}
# county-level, in most recent year
g <- bea %>% filter(GEOID %in% counties$GEOID, YEAR == 2019) %>% mutate(contrib = current_GDP_ag / current_GDP * 100) %>% filter(!is.na(contrib))

#464 counties > 10%

g_sf <- merge(counties, g, by = "GEOID")

# Pull quantiles
q<- g_sf %>%
  pull(contrib) %>%
  quantile(probs = seq(0, 1, length.out = 7), na.rm = T) %>%
  as.vector() 

quantiles <- c(0,0.5,1,2.5,5,15,30,90)
labels <- c("0 - 0.5%",      "0.5 - 1%",    "1 - 2.5%",    "2.5 - 5%",   "5 - 15%", "15 - 30%",  "30 - 80%")

# Map ag contribution to GDP
g_sf %>%
  mutate(quantiles = cut(contrib,
                               breaks = quantiles,
                               labels = labels,
                               include.lowest = T)) %>% 
  filter(YEAR == 2019) %>% 
ggplot(.) +
  geom_sf(aes(fill = quantiles), color = "transparent") +
  geom_sf(data = states, fill = "transparent", color = "dark gray") +
  theme_map() +
    theme(panel.grid.major = element_line(colour = "white"), 
          legend.position="right",
          legend.key.size = unit(4, 'mm')) +
  scale_fill_manual(values=c("#fef6b5","#ffdd9a","#ffc285","#ffa679","#fa8a76","#f16d7a","#e15383"),name="") 
  
  ggsave('./viz/SI-Figure1.png', dpi = 400, width = 110, height = 60, units = "mm", bg = "white")
```
### **SI Figure 2**: 

#### The proportion of all workers engaged in the direct production of agricultural commodities, livestock or crops (includes sole proprietors, partners, and hired laborers) in 2019. Data: US BEA
```{r}
# Farm employment as percentage of total county employment
# county-level, in most recent year (sf/ggplot())
fc <- bea %>% filter(GEOID %in% counties$GEOID) %>% select(c(1:2,38:39,44:45)) 
fc$farmempl <- as.numeric(apply(fc[c(5:6)],1,paste_noNA))
fc$totjobs <- as.numeric(apply(fc[c(3:4)],1,paste_noNA))

fc <- fc %>% mutate(prop_farm = farmempl/totjobs*100) %>% select(1:2,7:9)

fc_sf <- merge(counties, fc, by = "GEOID")

# Pull quantiles
q<- fc_sf %>%
  pull(prop_farm) %>%
  quantile(probs = seq(0, 1, length.out = 8), na.rm = T) %>%
  as.vector() 

quantiles <- c(0,2.5,5,7.5,10,15,30,75)
labels <- c("0 – 2.5%",      "2.5 - 5.0%",    "5.0 - 7.5%",    "7.5 - 10.0%",   "10.0 - 15.0%",  "15.0 - 30.0%", "> 30.0%")

# Map ag contribution to GDP
fc_sf %>%
  mutate(quantiles = cut(prop_farm,
                               breaks = quantiles,
                               labels = labels,
                               include.lowest = T)) %>% 
  filter(YEAR == 2019) %>%
  filter(!is.na(prop_farm)) %>% 
ggplot(.) +
  geom_sf(aes(fill = quantiles), color = "transparent") +
  geom_sf(data = states, fill = "transparent", color = "dark gray") +
  theme_map() +
    theme(panel.grid.major = element_line(colour = "white"), 
          legend.position="right",
          legend.key.size = unit(4, 'mm')) +
  scale_fill_manual(values=c("#f9ddda","#f2b9c4","#e597b9","#ce78b3","#ad5fad","#834ba0","#573b88"),name="") 

  ggsave('./viz/SI-Figure2.png', dpi = 400, width = 110, height = 60, units = "mm", bg = "white")
```

### **SI Figure 3**: 

#### (A.) Average age of primary operators in 2017. (B.) Change in average age between 1997 and 2017; where red, average age was less in 2017 than 1997. Data: USDA NASS QuickStats.
```{r}
############################################################
# Average Age
############################################################
age <- c_sf %>% select(c(1,9:13)) 

### Extract quantiles; use these to estimate quantiles, below!
q <- age %>% filter(year == 2017) %>% 
  pull(age) %>%
  quantile(probs = seq(0, 1, length.out = 7), na.rm = T) %>%
  as.vector() # to remove names of quantiles, so idx below is numeric

### What are the quantiles?
quantiles <- c(32.5,55,57.5,59,61.5,63,75)

### Create custom labels
labels <- c("32.5 – 55 years",      "55 – 57.5 years",    "57.5 – 59 years",    "59 – 61.5 years",   "61.5 – 63 years",  "> 63 years")

### Create a new variable with the quantiles
df <- age %>%
  mutate(mean_quantiles = cut(age,
                               breaks = quantiles,
                               labels = labels,
                               include.lowest = T)) %>% 
  filter(!is.na(age)) 

### PLOT
# Average age
ave_age <- ggplot(data = df) +
  geom_sf(mapping = aes(fill = mean_quantiles),color = "transparent",size = 0.01) +
  geom_sf(data = states, fill = "transparent", color = "dark gray") +
  scale_fill_brewer(palette = "Reds") +
  theme_map() +
    theme(panel.grid.major = element_line(colour = "white"), 
          legend.position="right",
          legend.key.size = unit(4, 'mm'))

############################################################
# Change in average Age between 1997 & 2017
############################################################
age_diff <- c %>% select(c(1:5)) %>%   
  filter(year %in% c(1997,2017)) %>% 
  group_by(GEOID, county) %>%
  summarise(age = diff(age)) %>% 
  filter(!is.na(age))

# percent cty's reporting reduction in average age
z<- age_diff %>% filter(age < 0) #67/3108 counties

t <- age_diff %>% filter(age <= 0)

age_diff <- merge(counties, age_diff, by = "GEOID")

### Extract quantiles; use these to estimate quantiles, below!
q<- age_diff %>%
  pull(age) %>%
  quantile(probs = seq(0, 1, length.out = 8), na.rm = T) %>%
  as.vector() # to remove names of quantiles, so idx below is numeric

quantiles <- c(-15,0,2.5,3.5,4.5,5.5,7.5,25)

labels <- c("< 0 years",      "0 - 2.5 years",    "2.5 - 3.5 years",    "3.5 - 4.5 years",   "4.5 - 5.5 years",  "5.5 - 7.5 years", "> 7.5 years")

age_diff <- age_diff %>%
  mutate(quantiles = cut(age,
                               breaks = quantiles,
                               labels = labels,
                               include.lowest = T)) %>% 
  filter(!is.na(age)) 

age_diff <- ggplot() +
  geom_sf(data = age_diff, aes(fill = quantiles), color = "transparent") +
  geom_sf(data = states, fill = "transparent", color = "dark gray") +
  scale_fill_manual(values = c("#B2182B","#F7F7F7", "#D1E5F0", "#92C5DE", "#4393C3", "#2166AC", "#053061")) +
  theme_map() +
    theme(panel.grid.major = element_line(colour = "white"), 
          legend.position="right",
          legend.key.size = unit(4, 'mm'))

# Build panel plots!
# (A.) % full owners, (B.) % tenants
library(ggpubr)
ggarrange(ave_age, age_diff, labels = c("A.", "B."), ncol = 2, font.label = list(size = 8))
ggsave("./viz/SI-Figure3.png", dpi = 400, units = "mm", width = 220, height = 60, bg = "white")
```

### **SI Figure 4**: 

#### Experience, years on present farm; The vertical line in 1997 reflects USDA-NASS implementation of adjustments for under-coverage. We do not include the non-adjusted ‘97 values. Data: Robert Hoppe, USDA ERS.
```{r}
# Experience

h %>%
  select(c(1,18:20)) %>% filter(!is.na(exp_05)) %>% mutate(perc = exp_10/(exp_05+exp_09+exp_10)) %>%
  pivot_longer(!c(1,5), names_to = "exp", values_to = "count") %>%  
ggplot(.) +
  geom_bar(aes(fill = exp, y = count, x = year), position = "fill", stat = "identity") +
  scale_fill_manual(name = "", labels = c("Less than 5", "5 to 9", "Greater than 10"), values=c("#117733","#44AA99","#88CCEE")) +
  labs(title = "", color = "", y = "", x = "") +
  geom_vline(xintercept=c(1997), linetype="dotted") +
  project_theme +
  theme(legend.position="bottom") 

  ggsave('./viz/SI-Figure4.png', dpi = 400, width = 100, height = 80, units = "mm")

# percent in 1910
1802540/5794768
# percent in 2017
2435435/3399834
```

### **SI Figure 5**: 

#### Primary operators reported years’ experience on present operation in 2017. Data: USDA NASS QuickStats.
```{r}
############################################################
# Years on any operation
############################################################

exp <- c_sf %>% select(c(1,10,12,17)) 

### Extract quantiles; use these to estimate quantiles, below!
q <- exp %>% filter(year == 2017) %>% 
  pull(exp) %>%
  quantile(probs = seq(0, 1, length.out = 7), na.rm = T) %>%
  as.vector() # to remove names of quantiles, so idx below is numeric

### What are the quantiles?
quantiles <- c(5,15,20,22.5,25,27.5,32.6)

### Create custom labels
labels <- c("5 - 15 years",      "15 - 20 years",    "20 – 22.5 years",    "22.5 - 25 years",   "25 - 27.5 years",  "> 27.5 years")

### Create a new variable with the quantiles
df <- exp %>%
  mutate(mean_quantiles = cut(exp,
                               breaks = quantiles,
                               labels = labels,
                               include.lowest = T)) %>% 
  filter(!is.na(exp)) %>% 
  filter(year == 2017) 

### PLOT
ggplot(data = df) +
  geom_sf(mapping = aes(fill = mean_quantiles),color = "transparent",size = 0.01) +
  geom_sf(data = states, fill = "transparent", color = "dark gray") +
  scale_fill_brewer(palette = "Blues") +
  theme_map() +
    theme(panel.grid.major = element_line(colour = "white"), 
          legend.position="right",
          legend.key.size = unit(4, 'mm'))

  ggsave('./viz/SI-Figure5.png', dpi = 400, width = 110, height = 60, units = "mm", bg = "white")
  
# STATS
z<- c %>% select(c(1,2,9)) %>% filter(year == 2017) %>% filter(exp >= 10) # (3058/3117)
```

### **SI Figure 6**: 

American Indian or Alaska Native .......................................46,210
Asian ...................................................................................16,978
Black or African American ..................................................38,447
Native Hawaiian or Other Pacific Islander ...........................2,306
White ...................................................................................2,614,526
More than one race reported ...............................................21,986

Hispanic principal producers' reported race:::
American Indian or Alaska Native .......................................2,077
Asian ...................................................................................861
Black or African American ...................................................879
Native Hawaiian or Other Pacific Islander ...........................439
White ...................................................................................84,434
More than one race reported ...............................................1,654

Principle Producers by race & hispanic origin (all races)
American Indian or Alaska Native .......................................44133
Asian ...................................................................................16117
Black or African American ...................................................37568
Native Hawaiian or Other Pacific Islander ...........................1867
White ...................................................................................2530092
More than one race reported ...............................................20332
Hispanic (all races) ...............................................90344

#### Proportional bar charts showing (A.) the area operated by principal operators by their reported race, and (B.) operations by principal operator’s race; y-axis limits at proportion = 0.08 to show distinctions between races (ie. white operators represent over 90% of all operators and farm over 90% of all operated acres). Data: USDA NASS QuickStats.
```{r}
re <- read.csv("./data/race-ethnicity.csv")

re$RACE <- factor(re$RACE, levels = c("White", "More", "Hawaiian", "Black", "Asian", "American Indian", "Hispanic"))

# Stacked bar chart, race
# area operated
race_area <- re %>%
  filter(TYPE == "Area Operated") %>% 
ggplot(.) +
  geom_bar(aes(fill = RACE, y = VALUE, x = YEAR), position = "fill", stat = "identity") +
  scale_fill_manual(name = "", labels = c("White", "Multiple Races", "Native Hawaiian or Other Pacific Islander", "Black", "Asian", "American Indian or Alaskan Native", "Hispanic"), values=c("#332288","#117733","#CC6677", "#44AA99","#88CCEE","#DDCC77","#882255")) +
  labs(title = "", color = "", y = "Proportion of Area Operated", x = "") +
  project_theme +
  theme(legend.position="bottom") +
  coord_cartesian(ylim = c(0,0.080)) +
  scale_x_continuous(breaks = c(2007, 2012, 2017)) 

# operations
race_operations <- re %>%
  filter(TYPE == "Operations") %>% 
ggplot(.) +
  geom_bar(aes(fill = RACE, y = VALUE, x = YEAR), position = "fill", stat = "identity") +
  scale_fill_manual(name = "", labels = c("White", "Multiple Races", "Native Hawaiian or Other Pacific Islander", "Black", "Asian", "American Indian or Alaskan Native", "Hispanic"), values=c("#332288","#117733","#CC6677", "#44AA99","#88CCEE","#DDCC77","#882255")) +
  labs(title = "", color = "", y = "Proportion of Operations", x = "") +
  #geom_vline(xintercept=c(1997), linetype="dotted") +
  project_theme +
  theme(legend.position="bottom") +
  coord_cartesian(ylim = c(0,0.080)) +
  scale_x_continuous(breaks = c(2007, 2012, 2017)) 

# operators
race_operators <- re %>%
  filter(TYPE == "Operators") %>% 
ggplot(.) +
  geom_bar(aes(fill = RACE, y = VALUE, x = YEAR), position = "fill", stat = "identity") +
  scale_fill_manual(name = "", labels = c("White", "Multiple Races", "Native Hawaiian or Other Pacific Islander", "Black", "Asian", "American Indian or Alaskan Native", "Hispanic"), values=c("#332288","#117733","#CC6677", "#44AA99","#88CCEE","#DDCC77","#882255")) +
  labs(title = "", color = "", y = "Proportion of Operators", x = "") +
  #geom_vline(xintercept=c(1997), linetype="dotted") +
  project_theme +
  theme(legend.position="bottom") +
  coord_cartesian(ylim = c(0,0.080)) +
  scale_x_continuous(breaks = c(2007, 2012, 2017)) 


# Build panel plots!
ggarrange(race_operators, race_area, race_operations, labels = c("A.", "B.", "C."), ncol = 1, nrow = 3, common.legend = TRUE, legend = "bottom",
          font.label = list(size = 8))
ggsave("./viz/SI-Figure6.png", dpi = 400, width = 150, height = 220, units = "mm", bg = "white")
```

### **SI Figure 7**: 

#### Percent acreage operated by (A.) Black, (B.) Hispanic, (C.) white, and (D.) female identifying primary operators in 2017. Data: USDA QuickStats, 2017 COA.
```{r}
############################################################
# Percent acres operated by principal operators identifying as females
############################################################
female_acres <- c_sf %>% select(c(1,10:11,21)) 

### Extract quantiles; use these to estimate quantiles, below!
q <- female_acres %>% filter(year == 2017) %>% 
  pull(percf_acr) %>%
  quantile(probs = seq(0, 1, length.out = 7), na.rm = T) %>%
  as.vector() # to remove names of quantiles, so idx below is numeric

### What are the quantiles?
quantiles <- c(0,15,20,22.5,25,30,77)

### Create custom labels
labels <- c("0 – 15%",      "15 – 20%",    "20 – 22.5%",    "22.5 – 25.0%",   "25.0 – 30.0%",  "> 30.0%")

### Create a new variable with the quantiles
df <- female_acres %>%
  mutate(mean_quantiles = cut(percf_acr,
                               breaks = quantiles,
                               labels = labels,
                               include.lowest = T)) %>% 
  filter(!is.na(percf_acr)) %>% 
  filter(year == 2017)

### PLOT
# Percent acres operated by women - 2017
female_acres <- ggplot(data = df) +
  geom_sf(mapping = aes(fill = mean_quantiles),color = "transparent",size = 0.01) +
  geom_sf(data = states, fill = "transparent", color = "dark gray") +
  scale_fill_brewer(palette = "Greens") +
  theme_map() +
    theme(panel.grid.major = element_line(colour = "white"), 
          legend.position="right",
          legend.key.size = unit(4, 'mm'))

############################################################
# Percent acres operated by folx identifying as black, hispanic, and white principal operators
############################################################

# proporation of each race
race_prop <- r %>% mutate(TOTAL = rowSums(.[7:13])) %>% 
  mutate(POC = (NATIVE_AMER + BLACK + MULTI + HISPANIC + ASIAN + PACIFIC_ISL)/TOTAL*100,
         NATIVE_AMER = NATIVE_AMER/TOTAL*100,
         BLACK = BLACK/TOTAL*100,
         MULTI = MULTI/TOTAL*100,
         WHITE = WHITE/TOTAL*100,
         HISPANIC = HISPANIC/TOTAL*100,
         ASIAN = ASIAN/TOTAL*100,
         PACIFIC_ISL = PACIFIC_ISL/TOTAL*100) %>% 
  filter(OPERATOR == "OPERATORS, PRINCIPAL") %>% 
  select("GEOID", "YEAR", "OPERATOR", "TYPE", "WHITE", "POC", "NATIVE_AMER", "BLACK", "MULTI", "HISPANIC", "ASIAN", "PACIFIC_ISL", "TOTAL") %>% 
  mutate_at(7:13, round, 3)

# build spatial
r_sf <- merge(counties, race_prop, by = "GEOID") 

### Extract quantiles; use these to estimate quantiles, below!
q <- r_sf %>% filter(YEAR == 2017, TYPE == "AREA OPERATED") %>% 
  pull(WHITE) %>%
  quantile(probs = seq(0, 1, length.out = 7), na.rm = T) %>%
  as.vector() 

### What are the quantiles?
q_h <- c(0,0.25,0.5,1.0,2.5,5.0,101)
q_b <- c(0,0.25,0.5,1.0,2.5,5.0,72)
q_w <- c(0,90,95,97.5,98.5,99.5,100.1)

### Create custom labels
l_h <- c("0 - 0.25%",      "0.25 - 0.5%",    "0.5 - 1.0%",    "1.0 - 2.5%",   "2.5 - 5.0%",  "> 5.0%")
l_b <- c("0 - 0.25%",      "0.25 - 0.5%",    "0.5 - 1.0%",    "1.0 - 2.5%",   "2.5 - 5.0%",  "> 5.0%")
l_w <- c("0 - 90.0%",    "90.0 - 95.0%",   "95.0 - 97.5%",   "97.5 - 98.5%", "98.5 - 99.5%",  "> 99.5%")

### Create a new variable with the quantiles
df <- r_sf %>%
  mutate(mean_quantiles = cut(HISPANIC,
                               breaks = q_h,
                               labels = l_h,
                               include.lowest = T)) %>% 
  filter(!is.na(HISPANIC)) %>% 
  filter(YEAR == 2017, TYPE == "AREA OPERATED") 

### PLOT
# Percent HISPANIC operators - 2017
hisp_acres <- ggplot(data = df) +
  geom_sf(mapping = aes(fill = mean_quantiles),color = "transparent",size = 0.01) +
  geom_sf(data = states, fill = "transparent", color = "dark gray") +
  scale_fill_brewer(palette = "Purples") +
  theme_map() +
    theme(panel.grid.major = element_line(colour = "white"), 
          legend.position="right",
          legend.key.size = unit(4, 'mm'))

### Create a new variable with the quantiles
df <- r_sf %>%
  mutate(mean_quantiles = cut(BLACK,
                               breaks = q_b,
                               labels = l_b,
                               include.lowest = T)) %>% 
  filter(!is.na(BLACK)) %>% 
  filter(YEAR == 2017, TYPE == "AREA OPERATED") 

### PLOT
# Percent BLACK operators - 2017
black_acres <- ggplot(data = df) +
  geom_sf(mapping = aes(fill = mean_quantiles),color = "transparent",size = 0.01) +
  geom_sf(data = states, fill = "transparent", color = "dark gray") +
  scale_fill_brewer(palette = "Oranges") +
  theme_map() +
    theme(panel.grid.major = element_line(colour = "white"), 
          legend.position="right",
          legend.key.size = unit(4, 'mm'))

### Create a new variable with the quantiles
df <- r_sf %>%
  mutate(mean_quantiles = cut(WHITE,
                               breaks = q_w,
                               labels = l_w,
                               include.lowest = T)) %>% 
  filter(!is.na(WHITE)) %>% 
  filter(YEAR == 2017, TYPE == "AREA OPERATED") 

### PLOT
# Percent WHITE operators - 2017
white_acres <- ggplot(data = df) +
  geom_sf(mapping = aes(fill = mean_quantiles),color = "transparent",size = 0.01) +
  geom_sf(data = states, fill = "transparent", color = "dark gray") +
  scale_fill_brewer(palette = "Reds") +
  theme_map() +
    theme(panel.grid.major = element_line(colour = "white"), 
          legend.position="right",
          legend.key.size = unit(4, 'mm'))

# Build panel plots!
# (A.) Black, (B.) Hispanic, (C.) white and (D.) female operated acreage
library(ggpubr)
ggarrange(black_acres, hisp_acres, white_acres, female_acres, labels = c("A.", "B.", "C.", "D."), ncol = 2, nrow = 2, font.label = list(size = 8))
ggsave("./viz/SI-Figure7.png", dpi = 400, width = 220, height = 120, units = "mm", bg = "white")
```
### **SI Figure 8**: 

#### (A.) Percent agricultural acres operated by POC and Hispanic-identifying  (i.e., Black, American Indian or Native Alaskan, Multi-racial, Asian, Native Hawaiian or other Pacific Islander) principal operators in 2017, and (B.) Percent of all US farm operations operated by POC and Hispanic-identifying principal operators in 2017. Data: USDA NASS QuickStats
```{r}
############################################################
# Percent acres operated by POC principal operators
############################################################

### Extract quantiles; use these to estimate quantiles, below!
q <- r_sf %>% filter(YEAR == 2017, TYPE == "AREA OPERATED") %>% 
  pull(POC) %>%
  quantile(probs = seq(0, 1, length.out = 7), na.rm = T) %>%
  as.vector() # to remove names of quantiles, so idx below is numeric

### What are the quantiles?
quantiles <- c(0,0.125,0.75,2.5,5,25,100)

### Create custom labels
labels <- c("0 - 0.125%",      "0.125 - 0.75%",    "0.75 - 2.5%",    "2.5 - 5.0%",   "5.0 - 25.0%",  "> 25.0%")

### Create a new variable with the quantiles
df <- r_sf %>%
  mutate(mean_quantiles = cut(POC,
                               breaks = quantiles,
                               labels = labels,
                               include.lowest = T)) %>% 
  filter(!is.na(POC)) %>% 
  filter(YEAR == 2017, TYPE == "AREA OPERATED") 

### PLOT
# Percent POC operated acres - 2017
POC_acres <- ggplot(data = df) +
  geom_sf(mapping = aes(fill = mean_quantiles),color = "transparent",size = 0.01) +
  geom_sf(data = states, fill = "transparent", color = "dark gray") +
  scale_fill_brewer(palette = "GnBu") +
  theme_map() +
    theme(panel.grid.major = element_line(colour = "white"), 
          legend.position="right",
          legend.key.size = unit(4, 'mm'))

############################################################
# Percent operations operated by POC principle operators
############################################################

### Extract quantiles; use these to estimate quantiles, below!
q <- r_sf %>% filter(YEAR == 2017, TYPE == "OPERATIONS") %>% 
  pull(POC) %>%
  quantile(probs = seq(0, 1, length.out = 7), na.rm = T) %>%
  as.vector() # to remove names of quantiles, so idx below is numeric

### What are the quantiles?
quantiles <- c(0,1.25,2.5,5,10,25,100)

### Create custom labels
labels <- c("0 - 1.25%",      "1.25 - 2.5%",    "2.5 - 5.0%",    "5.0 - 10.0%",   "10.0 - 25.0%",  "> 25.0%")

### Create a new variable with the quantiles
df <- r_sf %>%
  mutate(mean_quantiles = cut(POC,
                               breaks = quantiles,
                               labels = labels,
                               include.lowest = T)) %>% 
  filter(!is.na(POC)) %>% 
  filter(YEAR == 2017, TYPE == "OPERATIONS") 

### PLOT
# Percent POC operated acres - 2017
POC_operations <- ggplot(data = df) +
  geom_sf(mapping = aes(fill = mean_quantiles),color = "transparent",size = 0.01) +
  geom_sf(data = states, fill = "transparent", color = "dark gray") +
  scale_fill_brewer(palette = "YlGn") +
  theme_map() +
    theme(panel.grid.major = element_line(colour = "white"), 
          legend.position="right",
          legend.key.size = unit(4, 'mm'))

# Build panel plots!
# (A.) Percent POC operated acres, (B.) Percent POC-run operations
library(ggpubr)
ggarrange(POC_acres, POC_operations, labels = c("A.", "B."), ncol = 2, font.label = list(size = 8))
ggsave("./viz/SI-Figure8.png", dpi = 400, units = "mm", width = 220, height = 60, bg = "white")
```

### **SI Figure 9**: 

#### Proportion of acres operated by tenancy status. Data: USDA NASS QuickStats.
```{r}
c %>%
  select(c(1:2,18:20)) %>% 
  pivot_longer(!c(1,2), names_to = "tenancy", values_to = "acres") %>% 
ggplot(.) +
  geom_bar(aes(fill = forcats::fct_rev(tenancy), y = acres, x = year), position = "fill", stat = "identity") +
  scale_fill_manual(name = "", labels = c("Tenant", "Part Owner", "Full Owner"), values=c("#44AA99","#88CCEE","#882255")) +
  labs(title = "", color = "", y = "", x = "") +
  project_theme +
  theme(legend.position="bottom") +
  scale_x_continuous(breaks = c(1997, 2002, 2007, 2012, 2017)) 
  ggsave("./viz/SI-Figure9.png", dpi = 400, width = 100, height = 80, units = "mm")

# STATS
# tenancy through time

z <- c %>%
  select(c(1:2,18:20)) %>% 
  filter(year == 2017)   

z[is.na(z)] <- 0 

z <- z %>% group_by(year) %>% 
  summarise(full_owner = sum(full_owner),
            part_owner = sum(part_owner),
            tenant = sum(tenant))

# proportion of acres operated by tenancy status
# full owner
294702960/879236316
# part owner
501792110/879236316
# tenant
82741246/879236316
```

### **SI Figure 10**: 

#### Average acres per operation in 2017, measured as the total number of agricultural acres in a county divided by the total number of operations. Data: USDA NASS QuickStats.
```{r}
############################################################
# Acres per operation, mean in 2017
############################################################
size <- c %>% select(c(1:4,6:7)) %>% filter(year == 2017)  %>% mutate(ave_apo = round(tot_acres/n_op))

size <- merge(counties, size, by = "GEOID")

### Extract quantiles; use these to estimate quantiles, below!
# Mean
q <- size %>% 
  pull(ave_apo) %>% # switch ave_apo
  quantile(probs = seq(0, 1, length.out = 7), na.rm = T) %>%
  as.vector() # to remove names of quantiles, so idx below is numeric

### What are the quantiles?
quantiles <- c(0,150,300,600,1200,2500,60000) # Mean

### Create custom labels
labels <- c("0 – 150 acres",      "150 – 300 acres",    "300 – 600 acres",    "600 – 1200 acres",   "1200 – 2500 acres",  "> 2500 acres") # Mean

### Create a new variable with the quantiles
# Mean
df <- size %>%
  mutate(mean_quantiles = cut(ave_apo,
                               breaks = quantiles,
                               labels = labels,
                               include.lowest = T)) %>% 
  filter(!is.na(ave_apo)) 

### PLOT
# Mean acres
ggplot(data = df) +
  geom_sf(mapping = aes(fill = mean_quantiles),color = "transparent",size = 0.01) +
  geom_sf(data = states, fill = "transparent", color = "dark gray") +
  scale_fill_brewer(palette = "YlGn") +
  theme_map() +
    theme(panel.grid.major = element_line(colour = "white"), 
          legend.position="right",
          legend.key.size = unit(4, 'mm'))

  ggsave('./viz/SI-Figure10.png', dpi = 400, width = 110, height = 60, units = "mm", bg = "white")

```

### **Figure 11**: 

#### Share of gross farm cash receipts spent on all agricultural inputs. During the height of the Farm Crisis of the 1980s, for instance, US farmers spent over 90% of their gross earnings (ie. earnings before subtracting production expenses) on inputs to agriculture. 
```{r}
bea$YEAR<-as.numeric(bea$YEAR) 
bea %>% select(c(1:3,11,23)) %>% filter(GeoName == "United States") %>% mutate(prod_exp = (prod_exp)/cash_income*100) %>% 
  ggplot(.) +
  geom_line(aes(x = YEAR, y = prod_exp), color = "#88CCEE", size = 1.5) +
  project_theme +
  labs(title = "", color = "", y = "Share of Earnings (%) Spent on Inputs", x = "") 

  ggsave("./viz/SI-Figure11.png", dpi = 400, width = 100, height = 80, units = "mm")
```

### **Figure 12**: 

#### Agricultural input expenditure as a proportion of gross farm cash receipts in 2017; where values are > 100% county input expenditures surpassed gross farm earnings. For example, in Chattahoochee, GA expenditure on inputs surpassed gross earnings for the county by just under 1000%. Data: US BEA. 
```{r}
# county-level, in most recent year
g <- bea %>% select(c(1:3,11,23)) %>% mutate(perc = prod_exp/cash_income*100) %>% filter(!is.na(perc),                                                                                   YEAR == 2019) 
g_sf <- merge(counties, g, by = "GEOID")

# Pull quantiles
q<- g_sf %>%
  pull(perc) %>%
  quantile(probs = seq(0, 1, length.out = 7), na.rm = T) %>%
  as.vector() 

quantiles <- c(0,70,85,100,125,150,200,1000)
labels <- c("0 - 70%",      "70 - 85%",    "85 - 100%",    "100 - 125%",   "125 - 150%",  "150 - 200%", "> 200%")

# Map ag contribution to GDP
g_sf %>%
  mutate(quantiles = cut(perc,
                               breaks = quantiles,
                               labels = labels,
                               include.lowest = T)) %>% 
ggplot(.) +
  geom_sf(aes(fill = quantiles), color = "transparent") +
  geom_sf(data = states, fill = "transparent", color = "dark gray") +
  scale_fill_manual(values=c("#053061", "#4393C3", "#92C5DE",  "#FDDBC7", "#D6604D", "#B2182B", "#67001F"),name="") +
  theme_map() +
    theme(panel.grid.major = element_line(colour = "white"), 
          legend.position="right",
          legend.key.size = unit(4, 'mm'))

  ggsave('./viz/SI-Figure12.png', dpi = 400, width = 110, height = 60, units = "mm", bg = "white")
```

### **SI Figure 15**: 

#### Agricultural real estate debt from 1969 to 2019, adjusted for inflation using the chain-type GDP deflator provided by the USDA ERS. Data: USDA ERS.
```{r}
# tidy data
library(viridis)
w <- read.csv("./data/ers-wealthstatdata.csv")
w <- w[c(1,2,4:8,11)]

debt <- w %>% 
  filter(VariableDescriptionPart1 %in% c("Real estate","Farm sector debt","Non-real estate"))

debt$variable <- paste(debt$VariableDescriptionPart1, debt$VariableDescriptionPart2, sep="_")

debt <- debt[c(1,2,6,8:9)]
unique(debt$variable)

debt <- debt %>% mutate(REAL_GDP = Amount / (ChainType_GDP_Deflator/100))

debt_tidy <- debt %>% pivot_wider(names_from = variable, values_from = Amount)

# VIZ
debt %>%
  filter(variable %in% c("Real estate_Commercial banks","Real estate_CCC storage and drying loans","Real estate_Farm Credit System","Real estate_Farm Service Agency","Real estate_Farmer Mac","Real estate_Individuals and others","Real estate_Life insurance companies"),
         Year >= 1960) %>%
ggplot(.) +
    geom_bar(aes(fill=variable, y=REAL_GDP*1000, x=Year), position="stack", stat="identity") +
    scale_fill_viridis(discrete = T, option = "magma",labels = c("CCC Loans", "Commercial Banks", "Farm Credit System","Farm Service Agency","Farmer Mac","Individuals & Others", "Life Insurance"),name="Holder") +
    project_theme +
  #scale_fill_discrete(labels = c("CCC Loans", "Commercial Banks", "Farm Credit System","Farm Service Agency","Farmer Mac","Individuals & Others", "Life Insurance"), values = c("#000004FF", "#2D1160FF", "#721F81FF", "#B63679FF", "#F1605DFF", "#FEAF77FF", "#FCFDBFFF")) +
    xlab("") +
  labs(title = "", y = "Real US Dollars") +
  scale_y_continuous(breaks=c(100000000000,200000000000,300000000000),
        labels=c("$100B","$200B","$300B")) +
  theme(legend.position="bottom") 

  ggsave("./viz/SI-Figure15.png", dpi = 400, width = 140, height = 80, units = "mm")
```

### **SI Figure 16**: 

#### Agricultural non-real estate debt from 1969 to 2019, adjusted for inflation using the chain-type GDP deflator provided by the USDA ERS. Data: USDA ERS.
```{r}
debt %>%
  filter(variable %in% c("Non-real estate_Farm Credit System","Non-real estate_Farm Service Agency","Non-real estate_Individuals and others","Non-real estate_Commerical banks"),
         Year >= 1960) %>%
ggplot(.) +
    geom_bar(aes(fill=variable, y=REAL_GDP*1000, x=Year), position="stack", stat="identity") +
    scale_fill_viridis(discrete = T, labels = c("Farm Credit System","Farm Service Agency","Individuals & Others", "Commercial Banks"), name = "Holder") +
    ggtitle("Proportion of Non-real Estate Debt: By holder") +
    project_theme +
  labs(title = "", y = "Real US Dollars") +
  scale_y_continuous(breaks=c(50000000000,100000000000,150000000000,200000000000),
        labels=c("$50B","$100B","$150B","$200B")) +
  theme(legend.position="bottom") 

  ggsave("./viz/SI-Figure16.png", dpi = 400, width = 140, height = 80, units = "mm")
```
